---
title: "Plasmid Manuscript"
author: "Max Cummins"
date: "2024-05-16"
output:
  html_document: 
    code_folding: hide
  pdf_document: default
---

```{r setup, include=FALSE}
#Read in our packages
library(tidyverse)
library(vroom)
library(ggplot2)
library(xml2)
library(ggrepel)
library(caret)
library(here)
library(scales)
library(plotly)
library(ComplexHeatmap)
library(grid)
library(ComplexUpset)
library(hues)
library(ggpubr)

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	root.dir = here::here()
)
#Define our not in function
`%nin%` <- Negate(`%in%`)
```

## Define our color scheme
```{r}
#Overwrite with different colours
source_cols <- c("Companion Animal" = "#59398d",
         "Environmental" = "#709b46",
         "Food" = "#c09f3d",
         "Human" = "#48c595",
         "Livestock" =  "#c26bbc",
         "Wild Animal" = "#b9553d")

#Create fake colour for header
phylogroup_header <- c(NA)
names(phylogroup_header) <- 'Phylogroups'

#Define our colours for our phylogroups
clermont_cols <- c('#ffe119', '#4363d8', '#f58231', '#dcbeff', '#800000', '#000075', '#9A6324', 'red', '#3cb44b',  'white', '#a9a9a9', 'black')

#Assign our color names for phylogroups
names(clermont_cols) <- c("A", "B1",  "B2", "C", "D", "E", "F", "G", "E or cladeI", "cladeI", "cladeIII", "cladeIV")
```


## Read in our tree and genomic and metadata dataset

```{r Merge_analysis_outputs, include=FALSE}

genometa <- vroom("delims/genometa_n5471.txt", show_col_types = FALSE)

#Read in our tree
tree <- ggtree::read.tree("analysis/RapidNJ_5471.nwk")

#Remove quotes from sample names
tree$tip.label <- gsub("'", '', tree$tip.label)
tree$tip.label <- gsub("\\.fa", "", tree$tip.label)

#Create a list of samples in our cohort
sample_names <- tree$tip.label %>% as.data.frame() %>% rename('name' = '.')

```

## Results

```{r}
## General stats

# Compute total strains sequenced
total_seq <- genometa %>% filter(grepl("AUSMDU|AusGEM", name)) %>% nrow() %>% format(big.mark = ",")

# Compute total strains collated
total_collate <- genometa %>% filter(!grepl("AUSMDU|AusGEM", name)) %>% nrow()  %>% format(big.mark = ",")

# Compute total strains under analysis
total_analysis <- genometa %>% nrow() %>% format(big.mark = ",")

## Phylogroups

#Compute genomes per Phylogroup
Phylogroup_table <- genometa %>% group_by(Consensus_phylogroup, Revised_Source_Niche) %>% tally() %>% pivot_wider(names_from = "Revised_Source_Niche", values_from = "n", values_fill = 0) %>% rowwise() %>% mutate(Total = sum(c_across(where(is.numeric)))) %>% mutate("Percentage" = percent(Total/sum(.$Total), accuracy = 0.1))

#Extract the count/perc of B2s
B2_perc <- Phylogroup_table %>% filter(Consensus_phylogroup == "B2") %>% pull(Percentage) %>% format(big.mark = ",")
B2_count <- Phylogroup_table %>% filter(Consensus_phylogroup == "B2") %>% pull(Total) %>% format(big.mark = ",")

#Extract count of others
other_phylogroup_count <- Phylogroup_table %>% filter(Consensus_phylogroup %in% c("cladeI", "cladeIII", "cladeIV", "Unknown", "E or clade I")) %>% pull(Total) %>% sum() %>% format(big.mark = ",")
```


# Analysis of ST frequencies

```{r}

## STs
#Generate table of ST counts
ST_table <- genometa %>% group_by(ST_new, Consensus_phylogroup) %>% tally(sort = T) %>% mutate("Percentage" = percent(n/sum(.$n), accuracy = 0.1))

#Compute ST count
ST_count <- ST_table %>% nrow()

#Compute novel ST count
novel_ST_count <- ST_table %>% filter(grepl("\\^\\^", ST_new)) %>% nrow()

#Compute known ST count
known_ST_count <- ST_table %>% filter(!grepl("\\^\\^", ST_new)) %>% nrow()

#Generate strings for frequencies, counts and phylogroups for top five STs
top5 <- ST_table %>% head(5) %>% mutate(text_col = paste0(ST_new, " (phylogroup ", Consensus_phylogroup, "; ", Percentage, " (", n,"/",sum(.$n), ")"))


#Extract text string
freq_string <-  top5 %>% pull(text_col) %>% paste0(collapse =",")

#Compute total and perc
top5_total <- top5 %>% pull(n) %>% sum()
top5_perc <- percent(top5_total/nrow(genometa), accuracy = 0.1)
top5_total <- top5_total %>% format(big.mark = ",")

#Compute counts of non-novel and non-top5
other_STs <- ST_table %>% filter(ST_new %nin% top5$ST_new) %>% filter(!grepl("\\^\\^", ST_new))

#Compute total and perc for others
other_STs_total <- other_STs %>% filter(n > 1) %>% pull(n) %>% sum()
other_STs_perc <- percent(other_STs_total/nrow(genometa), accuracy = 0.1)
other_STs_total <- other_STs_total %>% format(big.mark = ",")
min_count_other <- other_STs %>% filter(n > 1) %>% pull(n) %>% min()
max_count_other <- other_STs %>% filter(n > 1) %>% pull(n) %>% max()

#Compute novel ST count
singleton_total <- ST_table %>% filter(n == 1) %>% pull(n) %>% sum()
singleton_perc <- percent(singleton_total/nrow(genometa), accuracy = 0.1)
singleton_total <- singleton_total %>%  format(big.mark = ",")

```


## Figure 1

```{r, fig.width=8.7, fig.height=11.69, fig.cap='Figure 1 - Collection composition. This sunburst chart visualises the number and proportion of strains among a given source and Intestinal/Extraintestinal niche. Note: A total of 7 Companion Animal strains are unable to be categorised as either Intestinal or Extraintestinal due to missing metadata – data not shown above due to small sample size (n = XX)'}

# Group by Source and Intestinal/Extraintestinal niche
sunburst_df <- genometa %>% group_by(Revised_Source_Niche, Intestinal_or_Extraintestinal) %>% tally() %>% ungroup()

# Rename columns for clarity
sunburst_df <- sunburst_df %>% rename('parent' = Revised_Source_Niche, 'label' = Intestinal_or_Extraintestinal, 'values' = n)

# Summarise the data to get total values for each parent
sunburst_df <- sunburst_df %>% group_by(parent) %>% summarise(values = sum(values)) %>% rename('label' = parent) %>% mutate(parent = "Total") %>% mutate(label = paste0(label, " (", values, ")")) %>% bind_rows(sunburst_df)

# Add a row for the total number of observations
sunburst_df <- sunburst_df %>% add_row(label= 'Total', parent = '', values = nrow(genometa))

# Put the label in the format "label (values)"
sunburst_df <- sunburst_df %>% mutate(label = paste0(label, " (", values, ")")) %>% mutate(label = str_replace(label, "\\).*", "\\)"))

# Group by parent and mutate labels
sunburst_df <- sunburst_df %>% group_by(parent) %>% mutate(county = sum(values)) %>% mutate(parent = paste0(parent, " (", county, ")")) %>% mutate(parent = str_replace(parent, '^ \\(.*\\)$', '')) %>% mutate(parent = str_replace(parent, ' \\(', '\n(')) %>% mutate(label = str_replace(label, ' \\(', '\n('))

# More label changing
sunburst_df <- sunburst_df %>% mutate(Revised_Source_Niche = str_replace(parent, "\n.*", "")) %>% mutate(Revised_Source_Niche = ifelse(Revised_Source_Niche == "Total", "", Revised_Source_Niche))

# Manually set the source names
sunburst_df[1,5] <- "Companion Animal"
sunburst_df[2,5] <- "Environmental"
sunburst_df[3,5] <- "Food"
sunburst_df[4,5] <- "Human"
sunburst_df[5,5] <- "Livestock"
sunburst_df[6,5] <- "Wild Animal"
sunburst_df[nrow(sunburst_df),5] <- "Total"

#Source Colours
Source_cols <- c("Companion Animal" = "#59398d",
         "Environmental" = "#709b46",
         "Food" = "#c09f3d",
         "Human" = "#48c595",
         "Livestock" =  "#c26bbc",
         "Wild Animal" = "#b9553d",
         "Plasmid" =  "#000000")

# Add a column for source colours
sunburst_df <- sunburst_df %>% mutate(Source_col = Source_cols[Revised_Source_Niche])

# Set the last row's source colour to white for aesthetics
sunburst_df[nrow(sunburst_df),ncol(sunburst_df)] <- "white"

# Set seed for reproducibility
set.seed(1)

# Select relevant columns for the sunburst plot
sunburst_df <- sunburst_df %>% select(label, values, parent, Source_col)

# Plot
fig <- plot_ly(sunburst_df,
               labels = ~label,
               parents = ~parent,
               values = ~values,
  type = 'sunburst', branchvalues = 'total') %>%
        layout(colorway = ~Source_col)

# View Plot
fig 
```


```{r}
# Read in our mobtyper data
mobtyper_df <- vroom("analysis/mobtyper_results_concatenated.txt", show_col_types = FALSE)

# Read in our data on the mobtyper database
if (!exists("clusters")) {
    if (file.exists("delims/plasmid_cluster_data.txt.gz")) {
        # Unzip the file
        gunzip("delims/plasmid_cluster_data.txt.gz", remove = FALSE, overwrite = TRUE)
        clusters <- vroom("delims/plasmid_cluster_data.txt")
    } else {
        source("scripts/plasmid_data_aggregate.R")
        clusters <- vroom("delims/plasmid_cluster_data.txt")
    }
}

# Filter in our mobtyper data and rename the sample ID column to name
mobtyper_df <- mobtyper_df %>%
    rename('name' = sample_id) %>%
    mutate(name = str_replace(name, ":.*", "")) %>%
    filter(name %in% tree$tip.label)

detected_clusters <- clusters %>%
    filter(primary_cluster_id %in% unique(mobtyper_df$primary_cluster_id))

# Calculate entropy for each sequence type (ST)
ST_entropy <- genometa %>%
    group_by(ST_new) %>%
    add_count(name = 'Count_of_ST') %>%
    ungroup() %>%
    select(name, ST_new, Count_of_ST) %>%
    inner_join(mobtyper_df, by = "name") %>%
    group_by(ST_new, primary_cluster_id) %>%
    summarise(counts = n(), Count_of_ST) %>%
    unique() %>%
    group_by(ST_new, primary_cluster_id) %>%
    mutate(ST_cluster_prob = counts / Count_of_ST) %>%
    mutate(log_ST_cluster_prob = log(ST_cluster_prob)) %>%
    mutate(prod_ST_cluster_prob_and_log = ST_cluster_prob * log_ST_cluster_prob) %>%
    group_by(primary_cluster_id) %>%
    mutate(entropy = abs(sum(prod_ST_cluster_prob_and_log))) %>%
    mutate('ST_entropy' = entropy) %>%
    select(primary_cluster_id, ST_entropy) %>%
    unique()

# Calculate entropy for each phylogroup
Phylogroup_entropy <- genometa %>%
    group_by(Consensus_phylogroup) %>%
    add_count(name = 'Count_of_Phylogroup') %>%
    ungroup() %>%
    select(name, Consensus_phylogroup, Count_of_Phylogroup) %>%
    inner_join(mobtyper_df, by = "name") %>%
    group_by(Consensus_phylogroup, primary_cluster_id) %>%
    summarise(counts = n(), Count_of_Phylogroup) %>%
    unique() %>%
    group_by(Consensus_phylogroup, primary_cluster_id) %>%
    mutate(Phylogroup_cluster_prob = counts / Count_of_Phylogroup) %>%
    mutate(log_Phylogroup_cluster_prob = log(Phylogroup_cluster_prob)) %>%
    mutate(prod_Phylogroup_cluster_prob_and_log = Phylogroup_cluster_prob * log_Phylogroup_cluster_prob) %>%
    group_by(primary_cluster_id) %>%
    mutate(entropy = abs(sum(prod_Phylogroup_cluster_prob_and_log))) %>%
    mutate('Phylogroup_entropy' = entropy) %>%
    select(primary_cluster_id, Phylogroup_entropy) %>%
    unique()

# Calculate entropy for each source niche
Source_entropy <- genometa %>%
    group_by(Revised_Source_Niche) %>%
    add_count(name = 'Count_of_Revised_Source_Niche') %>%
    ungroup() %>%
    select(name, Revised_Source_Niche, Count_of_Revised_Source_Niche) %>%
    inner_join(mobtyper_df, by = "name") %>%
    group_by(Revised_Source_Niche, primary_cluster_id) %>%
    summarise(counts = n(), Count_of_Revised_Source_Niche) %>%
    unique() %>%
    group_by(Revised_Source_Niche, primary_cluster_id) %>%
    mutate(Revised_Source_Niche_cluster_prob = counts / Count_of_Revised_Source_Niche) %>%
    mutate(log_Revised_Source_Niche_cluster_prob = log(Revised_Source_Niche_cluster_prob)) %>%
    mutate(prod_Revised_Source_Niche_cluster_prob_and_log = Revised_Source_Niche_cluster_prob * log_Revised_Source_Niche_cluster_prob) %>%
    group_by(primary_cluster_id) %>%
    mutate(entropy = abs(sum(prod_Revised_Source_Niche_cluster_prob_and_log))) %>%
    mutate('Source_entropy' = entropy) %>%
    select(primary_cluster_id, Source_entropy) %>%
    unique()

# Combine detected clusters with entropy measures
detected_clusters_w_entropy <- detected_clusters %>%
    left_join(Phylogroup_entropy, by = "primary_cluster_id", relationship = "many-to-many") %>%
    left_join(ST_entropy, by = "primary_cluster_id", relationship = "many-to-many") %>%
    left_join(Source_entropy, by = "primary_cluster_id", relationship = "many-to-many") %>%
    select(sample_id, matches("entropy"), Plasmid_CIA_status, Plasmid_count_of_AMR_genes, Plasmid_VF_count, matches("^inc"), `IncF RST`, everything()) %>%
    mutate(Plasmid_CIA_status = replace_na(Plasmid_CIA_status, 0)) %>%
    mutate(Plasmid_count_of_AMR_genes = replace_na(Plasmid_count_of_AMR_genes, 0)) %>%
    mutate(Plasmid_VF_count = replace_na(Plasmid_VF_count, 0))

novel_and_plasmid_counts <- genometa %>%
        #Select columns of interest
        select(name, Revised_Source_Niche, Consensus_phylogroup) %>%
        #group by source to enable count
        group_by(Revised_Source_Niche) %>%
        #add source count
        add_count(name = "Source_Count") %>%
        #remove grouping
        ungroup() %>%
        #Bind to our mobtyper data
        full_join(mobtyper_df, by = "name") %>%
        #Group by sample
        group_by(name) %>%
        #Count the number of plasmids in a sample
        mutate(count_plasmids = n()) %>%
        #Correct where a count of zero is misidentified as a count of 1 for samples with no mobtyper data
        mutate(count_plasmids = if_else(is.na(md5) & count_plasmids == 1, 0, count_plasmids)) %>%
        #Add a column to indicate whether a particular plasmid is novel or not
        mutate(is_novel =  if_else(grepl("novel", primary_cluster_id), 1, 0)) %>%
        #Correct where a gene is indicated as having a non novel plasmid but actually has no plasmids
        mutate(is_novel = if_else(is.na(md5) & is_novel == 0, NA, is_novel)) %>%
        #Remove grouping
        ungroup()

#Compute total count of plasmids
total_plasmids <- novel_and_plasmid_counts %>% filter(!is.na(primary_cluster_id)) %>% nrow()

#Compute total count of novel plasmids
novel_plasmids <- novel_and_plasmid_counts %>% filter(is_novel == 1) %>% nrow()

#Compute total count of clustered plasmids
clustered_plasmids <- novel_and_plasmid_counts %>% filter(is_novel == 0) %>% nrow()

#Compute total count of detected plasmid clusters
plasmid_clusters <- novel_and_plasmid_counts %>% filter(is_novel == 0) %>% select(primary_cluster_id) %>% unique() %>% nrow()

#Compute percent of plasmids which were clustered
perc_clustered <- percent(clustered_plasmids/total_plasmids, accuracy = 0.01)

#Compute percent of plasmids which were novel
perc_novel <- percent(clustered_plasmids/total_plasmids, accuracy = 0.01)

#Compute count of top 20 primary clusters
count_top_20 <- novel_and_plasmid_counts %>% mutate(primary_cluster_id_top20 = fct_lump_n(f = primary_cluster_id, n = 20, other_level = "Other")) %>% group_by(primary_cluster_id_top20) %>% tally() %>% filter(primary_cluster_id_top20 != "Other") %>% pull(n) %>% sum()

#Compute percentage of top 20 primary clusters
perc_top_20 <- percent(count_top_20/total_plasmids, accuracy = 0.1)

#Format count for top 20
count_top_20 <- format(count_top_20, big.mark = ",")

#Format count for total plasmids
total_plasmids <- format(total_plasmids, big.mark = ",")

#Generate table of Rep types detected
rep_table <- novel_and_plasmid_counts %>% filter(`rep_type(s)` != "-") %>% separate_rows(`rep_type(s)`, sep = ",") %>% select(`rep_type(s)`) %>% unique() %>% filter(!is.na(`rep_type(s)`)) %>% group_by(grepl("Inc", `rep_type(s)`), grepl("^rep_", `rep_type(s)`), grepl("Col", `rep_type(s)`)) %>% tally() %>% rename("Inc-type" = `grepl("Inc", \`rep_type(s)\`)`, "Col-type" = `grepl("Col", \`rep_type(s)\`)`, "Other type" = `grepl("^rep_", \`rep_type(s)\`)`)

count_Incs <- rep_table %>% filter(`Inc-type` == TRUE) %>% pull(n)
count_other_reps <- rep_table %>% filter(`Other type` == TRUE) %>% pull(n)
count_cols <- rep_table %>% filter(`Col-type` == TRUE) %>% pull(n)
```

# Table 1

```{r}
#Create the base of our table
plasmid_df <- novel_and_plasmid_counts %>%
  filter(!is.na(primary_cluster_id)) %>%
  mutate(primary_cluster_id_top20 = fct_lump_n(f = primary_cluster_id, n = 20, other_level = "Other")) %>%
  mutate(primary_cluster_id_top20 = if_else(str_detect(string = primary_cluster_id, pattern ="novel"), "Novel", primary_cluster_id_top20)) 

base_table <- plasmid_df %>% group_by(primary_cluster_id_top20) %>% tally(sort = T)

base_table <- base_table %>% mutate(move_order = if_else(primary_cluster_id_top20 == "Other", 1, 0)) %>% arrange(move_order) %>% select(-move_order) %>% filter(!is.na(primary_cluster_id_top20))

# Calculate the total for each numeric column
totals <- base_table %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
  mutate(primary_cluster_id_top20 = "Total")

# Bind the totals row to the original data frame
base_table <- bind_rows(base_table, totals)

#Generate predominant replicons
reps_per_top20_clust <- plasmid_df %>% separate_rows(`rep_type(s)`, sep = ",") %>% group_by(primary_cluster_id_top20, `rep_type(s)`) %>% tally() %>% arrange(desc(n)) %>% slice_head(n = 1) %>% unique() %>% arrange(match(primary_cluster_id_top20, base_table$primary_cluster_id_top20)) %>% select(-n)

#Generate predominant IncF RSTs
IncF_RSTs_top_20_clust <- genometa %>% select(name, Revised_Source_Niche, `IncF_RST..IncF RST`) %>% left_join(plasmid_df, by = 'name') %>% select(name, primary_cluster_id_top20, `IncF_RST..IncF RST`) %>% group_by(primary_cluster_id_top20, `IncF_RST..IncF RST`) %>% tally(sort = T) %>% slice_head(n = 1) %>% arrange(match(primary_cluster_id_top20, base_table$primary_cluster_id_top20)) %>% left_join(reps_per_top20_clust, by = "primary_cluster_id_top20") %>% filter(grepl("IncF",`rep_type(s)`)) %>% select(primary_cluster_id_top20, `IncF_RST..IncF RST`) %>% mutate(`IncF_RST..IncF RST` = replace_na(`IncF_RST..IncF RST`, "-"))

#Generate source-wise table
source_wise_primary_clusters <- plasmid_df %>% select(name, primary_cluster_id_top20, Revised_Source_Niche) %>% group_by(primary_cluster_id_top20, Revised_Source_Niche) %>% tally(sort = T) %>% pivot_wider(names_from = "Revised_Source_Niche", values_from = "n", values_fill = 0)

#Combine all our tables
table_plasmids <- left_join(base_table, reps_per_top20_clust, by = "primary_cluster_id_top20") %>%
  left_join(IncF_RSTs_top_20_clust, by = "primary_cluster_id_top20") %>%
  left_join(source_wise_primary_clusters, by = "primary_cluster_id_top20") %>%
  rename("Primary Cluster ID" = primary_cluster_id_top20,
         "Total" = n,
         "IncF RST" = `IncF_RST..IncF RST`,
        'Predominant Replicon'  = `rep_type(s)`) %>%
  select(!matches("Total"), Total) %>%
  mutate(Percentage = percent(Total/max(Total), accuracy = 0.01)) %>%
  mutate(order = case_when(`Primary Cluster ID` == "Other" ~ "3",
                           `Primary Cluster ID` == "Novel" ~ "2",
                           `Primary Cluster ID` == "Total" ~ "4",
                           .default = "1")) %>%
  arrange(order, desc(Total)) %>%
  mutate(`IncF RST` = replace_na(`IncF RST`, "-")) %>%
  select(-order)

knitr::kable(table_plasmids, align = "c")
```

# Statistical analysis for plasmids


```{r eval=FALSE, include=FALSE}
#Statistical analysis of novel plasmid carriage by source
anova_df_source_novel <- novel_and_plasmid_counts %>%
        select('name', 'Revised_Source_Niche','count_plasmids') %>%
        mutate(Revised_Source_Niche = factor(Revised_Source_Niche,
                                             levels = c("Food",
                                                        "Environmental",
                                                        "Livestock",
                                                        "Wild Animal",
                                                        "Human",
                                                        "Companion Animal"))) %>%
        unique()

set.seed(1)

anova_df_source_novel <- genometa %>% select(name, Consensus_phylogroup) %>% inner_join(anova_df_source_novel, by = 'name')

subsampled_anova_df_source_novel <- anova_df_source_novel %>% filter(Consensus_phylogroup %in% c("A", "B1", "B2", "C", "D", "E", "F", "G")) %>% group_by(Revised_Source_Niche) %>% add_count() %>% ungroup() %>% group_by(Revised_Source_Niche, Consensus_phylogroup) %>% add_count() %>% mutate(nn = (1/nn)) %>% ungroup() %>% group_by(Revised_Source_Niche) %>% sample_n(size = min(.$n), weight = nn)

subsampled_anova_df_source_novel %>% select(name) %>% as.data.frame() %>% write_delim("delims/stat_subset_names.txt")

subsampled_anova_Source_novel <- aov(count_plasmids ~ Revised_Source_Niche, data = subsampled_anova_df_source_novel)

summary(subsampled_anova_Source_novel)[[1]]

means_df <- subsampled_anova_df_source_novel %>%
    group_by(Revised_Source_Niche) %>%
    summarise(mean_count = mean(count_plasmids))

subsampled_anova_df_source_novel <- subsampled_anova_df_source_novel %>%
    mutate(Revised_Source_Niche = factor(Revised_Source_Niche, 
                                         levels = means_df$Revised_Source_Niche[order(means_df$mean_count)]))

# Fit the ANCOVA model, controlling for the control variable
ancova_model <- aov(count_plasmids ~ Revised_Source_Niche + Consensus_phylogroup, data = subsampled_anova_df_source_novel)

# Get the summary of the ANCOVA model
ancova_summary <- summary(ancova_model)

# Print the ANCOVA summary to see the structure
print(ancova_summary)

tukey_result <- TukeyHSD(ancova_model)



tukey_result <- TukeyHSD(subsampled_anova_Source_novel)

tukey_df <- as.data.frame(tukey_result$`Revised_Source_Niche`)

tukey_df <- tukey_df %>% filter(`p adj` < 0.05)

tukey_df$pair <- rownames(tukey_df)

# Filter for significant comparisons (adjust the threshold as needed)
significant_comparisons <- tukey_df %>% filter(`p adj` < 0.05)

comparisons <- strsplit(as.character(significant_comparisons$pair), "-")

sorted_comparisons <- lapply(comparisons, function(x) sort(x))
sorted_comparisons <- lapply(sorted_comparisons, function(x) paste(x, collapse = "-"))
sorted_comparisons <- sorted_comparisons[order(sapply(sorted_comparisons, `[`, 1))]

comparisons <- lapply(sorted_comparisons, function(x) strsplit(x, "-")[[1]])


max_y <- max(subsampled_anova_df_source_novel$count_plasmids)

p <- ggboxplot(subsampled_anova_df_source_novel, x = "Revised_Source_Niche", y = "count_plasmids", add = "jitter", palette = source_cols, fill = "grey70", color = "Revised_Source_Niche")

p1 <- p + geom_signif(
    comparisons = comparisons,
    map_signif_level = TRUE,
    y_position = seq(from = max_y + 1, by = 1, length.out = length(comparisons))
)  + theme(legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(linewidth = 0.5, linetype = 'solid', colour = "grey80"),
    panel.grid.minor = element_line(linewidth = 0.25, linetype = 'solid', colour = "grey90"),
    plot.title = element_text(hjust = 0.5)) +
  labs(x = "Source Niche", y = "Count of plasmids")
```


```{r}
#Statistical analysis of novel plasmid carriage by source
anova_df_source_novel <- novel_and_plasmid_counts %>%
        select('name', 'Revised_Source_Niche','count_plasmids') %>%
        mutate(Revised_Source_Niche = factor(Revised_Source_Niche,
                                             levels = c("Food",
                                                        "Environmental",
                                                        "Livestock",
                                                        "Wild Animal",
                                                        "Human",
                                                        "Companion Animal"))) %>%
        unique()

set.seed(1)

anova_df_source_novel <- genometa %>% select(name, Consensus_phylogroup) %>% inner_join(anova_df_source_novel, by = 'name')

subsampled_anova_df_source_novel <- anova_df_source_novel %>% filter(Consensus_phylogroup %in% c("A", "B1", "B2", "C", "D", "E", "F", "G")) %>% group_by(Revised_Source_Niche) %>% add_count() %>% ungroup() %>% group_by(Revised_Source_Niche, Consensus_phylogroup) %>% add_count() %>% mutate(nn = (1/nn)) %>% ungroup() %>% group_by(Revised_Source_Niche) %>% sample_n(size = min(.$n), weight = nn)

subsampled_anova_df_source_novel %>% select(name) %>% as.data.frame() %>% write_delim("delims/stat_subset_names.txt")

subsampled_anova_Source_novel <- aov(count_plasmids ~ Revised_Source_Niche, data = subsampled_anova_df_source_novel)

summary(subsampled_anova_Source_novel)[[1]]

means_df <- subsampled_anova_df_source_novel %>%
    group_by(Consensus_phylogroup) %>%
    summarise(mean_count = mean(count_plasmids))

subsampled_anova_df_source_novel <- subsampled_anova_df_source_novel %>%
    mutate(Consensus_phylogroup = factor(Consensus_phylogroup, 
                                         levels = means_df$Consensus_phylogroup[order(means_df$mean_count)]))

# Fit the ANCOVA model, controlling for the control variable
ancova_model <- aov(count_plasmids ~ Consensus_phylogroup + Revised_Source_Niche, data = subsampled_anova_df_source_novel)

# Get the summary of the ANCOVA model
ancova_summary <- summary(ancova_model)

# Print the ANCOVA summary to see the structure
print(ancova_summary)

tukey_result <- TukeyHSD(ancova_model)



tukey_result <- TukeyHSD(ancova_model)

tukey_df <- as.data.frame(tukey_result$Consensus_phylogroup)

tukey_df <- tukey_df %>% filter(`p adj` < 0.05)

tukey_df$pair <- rownames(tukey_df)

# Filter for significant comparisons (adjust the threshold as needed)
significant_comparisons <- tukey_df %>% filter(`p adj` < 0.05)

comparisons <- strsplit(as.character(significant_comparisons$pair), "-")

sorted_comparisons <- lapply(comparisons, function(x) sort(x))
sorted_comparisons <- lapply(sorted_comparisons, function(x) paste(x, collapse = "-"))
sorted_comparisons <- sorted_comparisons[order(sapply(sorted_comparisons, `[`, 1))]

comparisons <- lapply(sorted_comparisons, function(x) strsplit(x, "-")[[1]])


max_y <- max(subsampled_anova_df_source_novel$count_plasmids)

max_y <- max(subsampled_anova_df_source_novel$count_plasmids)

p <- ggboxplot(subsampled_anova_df_source_novel, x = "Consensus_phylogroup", y = "count_plasmids", add = "jitter", palette = clermont_cols, fill = "grey70", color = "Consensus_phylogroup")

p2 <- p + geom_signif(
    comparisons = comparisons,
    map_signif_level = TRUE,
    y_position = seq(from = max_y + 1, by = 1, length.out = length(comparisons))
) +  theme(legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(linewidth = 0.5, linetype = 'solid', colour = "grey80"),
    panel.grid.minor = element_line(linewidth = 0.25, linetype = 'solid', colour = "grey90"),
    plot.title = element_text(hjust = 0.5)) +
  labs(x = "Phylogroup", y = "Count of plasmids")
```


```{r eval=FALSE, include=FALSE}
# Fit the ANCOVA model, controlling for the control variable
ancova_model <- aov(count_plasmids ~ Revised_Source_Niche + Consensus_phylogroup, data = subsampled_anova_df_source_novel)

# Get the summary of the ANCOVA model
ancova_summary <- summary(ancova_model)

# Print the ANCOVA summary to see the structure
print(ancova_summary)

tukey_result <- TukeyHSD(ancova_model)


# Extract Tukey HSD results into a data frame
tukey_df <- as.data.frame(tukey_result$Consensus_phylogroup)

# Add comparison groups as separate columns
tukey_df <- tukey_df %>%
  rownames_to_column(var = "comparison") %>%
  separate(comparison, into = c("group1", "group2"), sep = "-")

# Plot using ggpubr
ggpubr::ggdotchart(tukey_df, x = "group1", y = "diff",
                   add = "segments",
                   color = "group2",
                   sorting = "ascending",
                   add.params = list(color = "blue"),
                   dot.size = 3) +
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2) +
  labs(title = "Tukey HSD Test Results",
       x = "Groups",
       y = "Difference in Means") +
  theme_pubr()




# Extract Tukey HSD results into a data frame
tukey_df <- as.data.frame(tukey_result$Revised_Source_Niche)

# Add comparison groups as separate columns
tukey_df <- tukey_df %>%
  rownames_to_column(var = "comparison") %>%
  separate(comparison, into = c("group1", "group2"), sep = "-")

ggboxplot(subsampled_anova_df_source_novel, x = "Consensus_phylogroup", y = "count_plasmids", add = "jitter", palette = clermont_cols, fill = "grey70", color = "Consensus_phylogroup")



table_plas_by_source <- subsampled_anova_df_source_novel %>% group_by(Revised_Source_Niche) %>% summarise(mean_plasmid_count = mean(count_plasmids))

table_plas_by_phylogroup <- subsampled_anova_df_source_novel %>% group_by(Consensus_phylogroup) %>% summarise(mean_plasmid_count = mean(count_plasmids))

```

# Novel plasmid frequency
```{r}
#aggregate novel plasmid data
agg_data <- novel_and_plasmid_counts %>%
  group_by(name, Revised_Source_Niche) %>%
  summarise(novel_count = sum(is_novel, na.rm = TRUE))

#Generate a table with source-wise novel plasmid carriage data
agg_data2 <- agg_data %>% group_by(Revised_Source_Niche) %>% summarise(Count = n(), mean = signif(mean(novel_count, accuracy = 0.1), 2), count_of_novels_per_source = sum(novel_count), count_of_novel_carrying_strains = sum(novel_count > 0))

#Compute total count of novels
count_novel_plasmids <- agg_data2$count_of_novels_per_source %>% sum()

#Isolate counts per source
human_count <- agg_data2 %>% filter(Revised_Source_Niche == "Human") %>% pull(Count)
companion_animal_count <- agg_data2 %>% filter(Revised_Source_Niche == "Companion Animal") %>% pull(Count)
food_count <- agg_data2 %>% filter(Revised_Source_Niche == "Food") %>% pull(Count)
livestock_count <- agg_data2 %>% filter(Revised_Source_Niche == "Livestock") %>% pull(Count)
wild_animal_count <- agg_data2 %>% filter(Revised_Source_Niche == "Wild Animal") %>% pull(Count)
environment_count <- agg_data2 %>% filter(Revised_Source_Niche == "Environmental") %>% pull(Count)

#Isolate mean of novel plasmids
human_mean <- agg_data2 %>% filter(Revised_Source_Niche == "Human") %>% pull(mean)
companion_animal_mean <- agg_data2 %>% filter(Revised_Source_Niche == "Companion Animal") %>% pull(mean)
food_mean <- agg_data2 %>% filter(Revised_Source_Niche == "Food") %>% pull(mean)
livestock_mean <- agg_data2 %>% filter(Revised_Source_Niche == "Livestock") %>% pull(mean)
wild_animal_mean <- agg_data2 %>% filter(Revised_Source_Niche == "Wild Animal") %>% pull(mean)
environment_mean <- agg_data2 %>% filter(Revised_Source_Niche == "Environmental") %>% pull(mean)

#Isolate sum of novel plasmids
human_novel_count <- agg_data2 %>% filter(Revised_Source_Niche == "Human") %>% pull(count_of_novels_per_source)
companion_animal_novel_count <- agg_data2 %>% filter(Revised_Source_Niche == "Companion Animal") %>% pull(count_of_novels_per_source)
food_novel_count <- agg_data2 %>% filter(Revised_Source_Niche == "Food") %>% pull(count_of_novels_per_source)
livestock_novel_count <- agg_data2 %>% filter(Revised_Source_Niche == "Livestock") %>% pull(count_of_novels_per_source)
wild_animal_novel_count <- agg_data2 %>% filter(Revised_Source_Niche == "Wild Animal") %>% pull(count_of_novels_per_source)
environment_novel_count <- agg_data2 %>% filter(Revised_Source_Niche == "Environmental") %>% pull(count_of_novels_per_source)

#Compute percent of novel plasmids
human_novel_perc <- percent(human_novel_count/count_novel_plasmids, accuracy = 0.1)
companion_animal_novel_perc <- percent(companion_animal_novel_count/count_novel_plasmids, accuracy = 0.1)
food_novel_perc <- percent(food_novel_count/count_novel_plasmids, accuracy = 0.1)
livestock_novel_perc <- percent(livestock_novel_count/count_novel_plasmids, accuracy = 0.1)
wild_animal_novel_perc <- percent(wild_animal_novel_count/count_novel_plasmids, accuracy = 0.1)
environment_novel_perc <- percent(environment_novel_count/count_novel_plasmids, accuracy = 0.1)

#Compute count of strains from a given source novel plasmids
human_novel_carrying_strains <- agg_data2 %>% filter(Revised_Source_Niche == "Human") %>% pull(count_of_novel_carrying_strains)
companion_animal_novel_carrying_strains <- agg_data2 %>% filter(Revised_Source_Niche == "Companion_Animal") %>% pull(count_of_novel_carrying_strains)
food_novel_carrying_strains <- agg_data2 %>% filter(Revised_Source_Niche == "Food") %>% pull(count_of_novel_carrying_strains)
livestock_novel_carrying_strains <- agg_data2 %>% filter(Revised_Source_Niche == "Livestock") %>% pull(count_of_novel_carrying_strains)
wild_animal_novel_carrying_strains <- agg_data2 %>% filter(Revised_Source_Niche == "Wild Animal") %>% pull(count_of_novel_carrying_strains)
environment_novel_carrying_strains <- agg_data2 %>% filter(Revised_Source_Niche == "Environmental") %>% pull(count_of_novel_carrying_strains)

#Compute percent of strains from a given source novel plasmids
human_novel_carrying_perc <- percent(human_novel_carrying_strains/human_count, accuracy = 0.1)
companion_animal_novel_carrying_perc <- percent(companion_animal_novel_carrying_strains/companion_animal_count, accuracy = 0.1)
food_novel_carrying_perc <- percent(food_novel_carrying_strains/food_count, accuracy = 0.1)
livestock_novel_carrying_perc <- percent(livestock_novel_carrying_strains/livestock_count, accuracy = 0.1)
wild_animal_novel_carrying_perc <- percent(wild_animal_novel_carrying_strains/wild_animal_count, accuracy = 0.1)
environment_novel_carrying_perc <- percent(environment_novel_carrying_strains/wild_animal_count, accuracy = 0.1)

#Reformat human count for inline
human_count <- human_count %>% format(big.mark = ",")


```

# F plasmid analyses

```{r}
mobtyper_predominant_reps <- mobtyper_df %>% separate_rows(`rep_type(s)`, sep = ",") %>% mutate(`rep_type(s)` = paste0("Rep_Gene_", `rep_type(s)`)) %>% mutate(val_col = 1) %>% group_by(primary_cluster_id)  %>% select(primary_cluster_id, `rep_type(s)`, val_col) %>% pivot_wider(names_from = "rep_type(s)", values_from = val_col, values_fill = 0, values_fn = sum) %>% select(-`Rep_Gene_-`) %>% pivot_longer(names_to = "rep_type(s)", cols = 2:ncol(.)) %>% arrange(desc(value)) %>% group_by(primary_cluster_id) %>% add_count() %>% slice_head(n = 1) %>% mutate(`rep_type(s)` = str_replace(`rep_type(s)`, "Rep_Gene_", ""))

mobtyper_df_w_entropy <- detected_clusters_w_entropy %>% select(primary_cluster_id, Phylogroup_entropy) %>% unique() %>% inner_join(mobtyper_predominant_reps)

mobtyper_df_w_entropy <- mobtyper_df %>% group_by(primary_cluster_id) %>% add_count() %>% select(primary_cluster_id, n) %>% inner_join(mobtyper_df_w_entropy, by = "primary_cluster_id")

mobtyper_df_w_entropy <- mobtyper_df_w_entropy %>% mutate(`rep_type(s)` = str_replace(`rep_type(s)`, "^rep_cluster.*", "Other"))

mobtyper_df_w_entropy <- mobtyper_df_w_entropy %>% mutate(`rep_type(s)` = str_replace(`rep_type(s)`, "IncI.*|IncK.*", "IncI/B/O/K/Z")) %>% mutate(`rep_type(s)` = str_replace(`rep_type(s)`, "IncF.*", "IncF")) %>% mutate(`rep_type(s)` = str_replace(`rep_type(s)`, "Col.*", "Col-type")) %>% mutate(`rep_type(s)` = str_replace(`rep_type(s)`, "IncX.*", "IncX")) %>% rename("Predominant Replicon" = "rep_type(s)") %>% rename("Count of Cluster" = n.x, "Phylogroup Entropy" = Phylogroup_entropy)

rep_cols <- iwanthue(length(unique(mobtyper_df_w_entropy$`Predominant Replicon`)))

names(rep_cols) <- sort(unique(mobtyper_df_w_entropy$`Predominant Replicon`))

mobtyper_df_w_entropy %>% ggplot() + geom_point(aes(x = `Phylogroup Entropy`, y = `Count of Cluster`, color = `Predominant Replicon`)) + scale_color_manual(values = rep_cols)
```


```{r}
count_total_clusters <- detected_clusters_w_entropy %>% select(primary_cluster_id) %>% unique() %>% nrow()

plasmids_by_AMR_and_VAG <- detected_clusters_w_entropy %>% group_by(primary_cluster_id) %>% summarise(Count_w_VF = sum(Plasmid_VF_count > 0), Count_Total = n(), proportion_Vir = Count_w_VF/Count_Total, Count_w_AMR = sum(Plasmid_count_of_AMR_genes > 0), proportion_AMR = Count_w_AMR/Count_Total)

count_AMR_clusters <- plasmids_by_AMR_and_VAG %>% filter(proportion_AMR > 0.6) %>% nrow()

count_VAG_clusters <- plasmids_by_AMR_and_VAG %>% filter(proportion_Vir > 0.6) %>% nrow()

count_VAG_and_AMR_clusters <- plasmids_by_AMR_and_VAG %>% filter(proportion_Vir > 0.6 & proportion_AMR > 0.6) %>% nrow()

perc_AMR_clusters <- percent(count_AMR_clusters/count_total_clusters, accuracy = 0.1)

perc_VAG_clusters <- percent(count_VAG_clusters/count_total_clusters, accuracy = 0.1)

plasmids_by_AMR_and_VAG %>% left_join(unique(mobtyper_df_w_entropy)) %>% rename("Proportion Carrying Virulence Genes" = `proportion_Vir`) %>% ggplot() + geom_point(aes(x = `Phylogroup Entropy`, y = `Proportion Carrying Virulence Genes`, color = `Predominant Replicon`)) + scale_color_manual(values = rep_cols)
```


```{r}
clusters_w_classification <- detected_clusters_w_entropy %>% separate_rows(`rep_type(s)`, sep = ",") %>% mutate(`rep_type(s)` = paste0("Rep_Gene_", `rep_type(s)`)) %>% mutate(val_col = 1) %>% group_by(primary_cluster_id) %>% pivot_wider(names_from = "rep_type(s)", values_from = val_col, values_fill = 0, values_fn = list(val_col = function(x) as.integer(sum(x) > 0))) %>% select(sample_id, primary_cluster_id, Plasmid_count_of_AMR_genes, Plasmid_VF_count, Plasmid_CIA_status, matches('Rep_Gene')) %>% group_by(primary_cluster_id) %>% mutate(Count_w_VF = sum(Plasmid_VF_count > 0), Count_Total = n(), proportion_Vir = Count_w_VF/Count_Total, Count_w_AMR = sum(Plasmid_count_of_AMR_genes > 0), proportion_AMR = Count_w_AMR/Count_Total)

clusters_w_classification %>% filter(proportion_AMR > 0.6) %>% group_by(primary_cluster_id) %>% summarise(across(matches("Rep_Gene"), sum, na.rm = TRUE), Count_Total = unique(Count_Total)) %>% group_by(primary_cluster_id) %>% summarise(across(matches("Rep_Gene"), ~ . / Count_Total)) %>% column_to_rownames("primary_cluster_id") %>% select_if(~ !(all(. == 0, na.rm = TRUE))) %>%
  rename_with(~ gsub("Rep_Gene_", "", .), everything())%>% pheatmap(fontsize_row = 4, colorRampPalette(c("white", "red"))(100))

clusters_w_classification %>% filter(proportion_Vir > 0.6) %>% group_by(primary_cluster_id) %>% summarise(across(matches("Rep_Gene"), sum, na.rm = TRUE), Count_Total = unique(Count_Total)) %>% group_by(primary_cluster_id) %>% summarise(across(matches("Rep_Gene"), ~ . / Count_Total)) %>% column_to_rownames("primary_cluster_id") %>% select_if(~ !(all(. == 0, na.rm = TRUE))) %>%
  rename_with(~ gsub("Rep_Gene_", "", .), everything()) %>% pheatmap(fontsize_row = 4, colorRampPalette(c("white", "red"))(100))


clusters_w_classification %>% filter(proportion_Vir > 0.6 & proportion_AMR > 0.6) %>% group_by(primary_cluster_id) %>% summarise(across(matches("Rep_Gene"), sum, na.rm = TRUE), Count_Total = unique(Count_Total)) %>% group_by(primary_cluster_id) %>% summarise(across(matches("Rep_Gene"), ~ . / Count_Total)) %>% column_to_rownames("primary_cluster_id") %>% select_if(~ !(all(. == 0, na.rm = TRUE))) %>%
  rename_with(~ gsub("Rep_Gene_", "", .), everything()) %>% pheatmap(fontsize_row = 8, colorRampPalette(c("white", "red"))(100))

```

ColV and senB-associated plasmids exhibit different phylogenetic and source distributions but are associated with virulence and resistance carriage (560 words)
We identified the 12 most frequent virulence associated plasmid clusters identified in our collection and analysed their distribution throughout our genome collection to determine their sourcewise prevalence and phylogenetic promiscuity (measured as phylogroup entropy) (Supplementary Figure X). Most commonly, these plasmid clusters carrying virulence loci were associated with ColV virulence loci or senB/cjrABC (Table 1).

```{r}
virulence_clusters <- clusters_w_classification %>% filter(proportion_Vir > 0.6) %>% select(primary_cluster_id) %>% unique()

base_table <- mobtyper_df_w_entropy %>% filter(primary_cluster_id %in% virulence_clusters$primary_cluster_id) %>% unique() %>% arrange(desc(`Count of Cluster`)) %>% select(primary_cluster_id, `Phylogroup Entropy`,  `Count of Cluster`)

#Add in predominant F type
base_table <- mobtyper_df %>% select(name, primary_cluster_id) %>% left_join(genometa, by = 'name') %>% select(name, primary_cluster_id, `IncF_RST..IncF RST`) %>% inner_join(base_table) %>% group_by(primary_cluster_id, `IncF_RST..IncF RST`) %>% add_count() %>% filter(`IncF_RST..IncF RST` != "F-:A-:B-") %>% arrange(desc(n)) %>% ungroup() %>% group_by(primary_cluster_id) %>% slice_head(n =1) %>% arrange(desc(`Count of Cluster`)) %>% select(-name, -n)

#Add in ColV or SenB carriage
base_table <- mobtyper_df %>% select(name, primary_cluster_id) %>% left_join(genometa, by = 'name') %>% select(name, primary_cluster_id, ColV, ABRICATE..vfdb_senB) %>% inner_join(base_table) %>% select(primary_cluster_id, ColV, ABRICATE..vfdb_senB) %>% group_by(primary_cluster_id) %>% 
    # Calculate the proportion of rows carrying each gene
    summarise(across(ColV:ABRICATE..vfdb_senB, ~ sum(. > 0, na.rm = TRUE) / n(), .names = "prop_{col}")) %>% inner_join(base_table, by = "primary_cluster_id") %>% arrange(desc(`Count of Cluster`))

base_table <- mobtyper_df %>% select(name, primary_cluster_id) %>% left_join(genometa, by = "name") %>% select(name, primary_cluster_id, Revised_Source_Niche) %>% group_by(primary_cluster_id, Revised_Source_Niche) %>% tally(sort = T) %>% pivot_wider(names_from = "Revised_Source_Niche", values_from = n, values_fill = 0) %>% inner_join(base_table, by = 'primary_cluster_id')

base_table <- base_table %>% rename("primary_cluster_id" = primary_cluster_id, "ColV+ (%)" = prop_ColV, "senB+ (%)" = prop_ABRICATE..vfdb_senB, "IncF RST" = `IncF_RST..IncF RST`) %>% mutate(`ColV+ (%)` = percent(`ColV+ (%)`, accuracy = 0.1), `senB+ (%)` = percent(`senB+ (%)`, accuracy = 0.1), `Phylogroup Entropy` = round(`Phylogroup Entropy`, digits = 3)) %>% select(`primary_cluster_id`, `IncF RST`, `ColV+ (%)`, `senB+ (%)`, `Phylogroup Entropy`, Human, Livestock, `Wild Animal`, Food, Environmental, `Companion Animal`, `Count of Cluster`) %>% arrange(desc(`Count of Cluster`)) %>% head(12)

base_table %>% knitr::kable(align = 'c')
```


```{r}
colV_carriers <- genometa %>% pull(ColV) %>% sum()

senB_carriers <- genometa %>% pull(ABRICATE..vfdb_senB) %>% sum()

combined_colV_senB <- (colV_carriers + senB_carriers) %>% format(big.mark = ",")

colV_carriers <- colV_carriers %>% format(big.mark = ",")

senB_carriers <- senB_carriers %>% format(big.mark = ",")


senB_and_ColV <- genometa %>% select(ColV, ABRICATE..vfdb_senB) %>%
    rowwise() %>%
    mutate(row_sum = sum(c_across(ColV:ABRICATE..vfdb_senB))) %>%
    ungroup() %>%  # It's good practice to ungroup after rowwise operations
    group_by(row_sum > 1) %>%
  tally() %>%
  filter(`row_sum > 1` == TRUE) %>%
  pull(n)

senB_and_ColV_ST <- genometa %>% select(ST_new, ColV, ABRICATE..vfdb_senB) %>%
    rowwise() %>%
    mutate(row_sum = sum(c_across(ColV:ABRICATE..vfdb_senB))) %>%
    ungroup() %>%  # It's good practice to ungroup after rowwise operations
    group_by(row_sum > 1, ST_new) %>%
  tally() %>%
  filter(`row_sum > 1` == TRUE) %>%
  arrange(desc(n)) %>% head(1) %>% mutate(ST_new = paste0("ST", ST_new)) %>% pull(ST_new)

senB_and_ColV_ST_count <- genometa %>% select(ST_new, ColV, ABRICATE..vfdb_senB) %>%
    rowwise() %>%
    mutate(row_sum = sum(c_across(ColV:ABRICATE..vfdb_senB))) %>%
    ungroup() %>%  # It's good practice to ungroup after rowwise operations
    group_by(row_sum > 1, ST_new) %>%
  tally() %>%
  filter(`row_sum > 1` == TRUE) %>%
  arrange(desc(n)) %>% head(1) %>% pull(n)
```



# Co-association of shared plasmid and cgMLST overlap

```{r}
# Conditionally open analysis/cgMLST_dists_5471.txt or unzip analysis/cgMLST_dists_5471.txt.gz
if (file.exists("analysis/cgMLST_dists_5471.txt")) {
  dists <- read_delim("analysis/cgMLST_dists_5471.txt", delim = "\t")
} else if (file.exists("analysis/cgMLST_dists_5471.txt.gz")) {
  cgMLST_dists_data_file <- unzip("analysis/cgMLST_dists_5471.txt.gz")
  dists <- read_delim(cgMLST_dists_data_file, delim = "\t")
} else {
  stop("Neither analysis/cgMLST_dists_5471.txt nor analysis/cgMLST_dists_5471.txt.gz exists. Pull the repo again!")
}

#Convert to long format and rename columns
dists <- dists %>% pivot_longer(names_to = "Sample 2", cols = 2:ncol(.)) %>% rename(`Sample 1` = `5471`) %>% filter(`Sample 1` != `Sample 2`)

#Bind in meta/genotypic data
dists <-  genometa %>% select(name, Revised_Source_Niche, ST_new, Collection_Year) %>% rename(`Sample 1` = name) %>% inner_join(dists, by = 'Sample 1')

#Bind in meta/genotypic data
dists <-  genometa %>% select(name, Revised_Source_Niche, ST_new, Collection_Year) %>% rename(`Sample 2` = name) %>% inner_join(dists, by = 'Sample 2')

#Filter to close genomes
close_dists <- dists %>% filter(value <= 10)

#Count total number of pairs of close genomes
count_close_genome_pairs <- nrow(close_dists) %>% format(big.mark = ',')

#Filter close genomes which differ in source of origin
close_dists_diff_source <- close_dists %>% filter(Revised_Source_Niche.x != Revised_Source_Niche.y)

#Filter close genomes which share a source of origin
close_dists_same_source <- close_dists %>% filter(Revised_Source_Niche.x == Revised_Source_Niche.y)

#Count total number of genomes which were close (≤ 10 cgMLST alleles)
count_less_equal_10 <- close_dists %>%
  select(`Sample 1`, `Sample 2`) %>%
  distinct() %>%
  pivot_longer(cols = everything(), values_to = "unique_value") %>%
  distinct(unique_value) %>% nrow() %>% format(big.mark = ',')

#Count total number of genomes which were close (≤ 10 cgMLST alleles) and differed in source
count_less_equal_10_diff_source <- close_dists_diff_source %>%
  select(`Sample 1`, `Sample 2`) %>%
  distinct() %>%
  pivot_longer(cols = everything(), values_to = "unique_value") %>%
  distinct(unique_value) %>% nrow()

#Count total number of genome pairs whichw ere close and differed in source
count_pairs_diff_source <- close_dists_diff_source %>% nrow()

#Count total number of genomes which were close (≤ 10 cgMLST alleles) and shared source
count_less_equal_10_same_source <- close_dists_same_source %>%
  select(`Sample 1`, `Sample 2`) %>%
  distinct() %>%
  pivot_longer(cols = everything(), values_to = "unique_value") %>%
  distinct(unique_value) %>% nrow() %>% format(big.mark = ',')

#Count total number of genome pairs whichw ere close and differed in source
count_pairs_same_source <- close_dists_same_source %>% nrow() %>% format(big.mark = ',')

#Bind in plasmid data
close_dists_w_plas <- mobtyper_df %>% select(primary_cluster_id, secondary_cluster_id, `rep_type(s)`, name) %>% rename(`Sample 1` = 'name') %>% right_join(close_dists_diff_source, by = 'Sample 1')

#Bind in plasmid data
close_dists_w_plas <- mobtyper_df %>% select(primary_cluster_id, secondary_cluster_id, `rep_type(s)`, name) %>% rename(`Sample 2` = 'name') %>% right_join(close_dists_w_plas, by = 'Sample 2')

#Filter to genomes which carry plasmids in our plasmid table (F plasmids)
close_dists_w_plas <- close_dists_w_plas %>% filter(primary_cluster_id.x %in% base_table$primary_cluster_id & primary_cluster_id.y %in% base_table$primary_cluster_id)

#Cont number of primary clusters in this cohort with F plasmids
count_f_plasmids_in_close <- close_dists_w_plas %>% select(primary_cluster_id.x, primary_cluster_id.y) %>%
  distinct() %>%
  pivot_longer(cols = everything(), values_to = "unique_value") %>%
  distinct(unique_value) %>% nrow()

#Cont number of isolates in this cohort with F plasmids
count_f_plasmids_in_close <- close_dists_w_plas %>%
  select(`Sample 1`, `Sample 2`) %>%
  distinct() %>%
  pivot_longer(cols = everything(), values_to = "unique_value") %>%
  distinct(unique_value) %>% nrow()

#Count pairs of strains which shared a primary plasmid cluster
count_close <- close_dists_w_plas %>% nrow()

#Count pairs of strains which shared a primary plasmid cluster
count_primary_plas_clusts_close <- close_dists_w_plas %>%
    filter(primary_cluster_id.x == primary_cluster_id.y) %>% nrow()

#Count pairs of strains which shared a secondary plasmid cluster
count_secondary_plas_clusts_close <- close_dists_w_plas %>%
    filter(secondary_cluster_id.x == secondary_cluster_id.y) %>% nrow()
```



```{r}

library(circlize)

# Create an adjacency matrix: 
chord_data <- close_dists_w_plas %>%
        filter(value != 0) %>% mutate("Source_Combo" = paste0(pmin(as.character(Revised_Source_Niche.x), as.character(Revised_Source_Niche.y)), "/", pmax(as.character(Revised_Source_Niche.x), as.character(Revised_Source_Niche.y)))) %>%
        group_by(primary_cluster_id.x, Source_Combo) %>% summarise(counts = n()) %>%
        pivot_wider(id_cols = primary_cluster_id.x, names_from = Source_Combo, values_from = counts, values_fill = 0) %>% column_to_rownames("primary_cluster_id.x") %>% t() %>% as.data.frame() %>% select(AA171, AA337, AA297, AA735, AA315, AA297, AA174, AE638, AA179, AA175, AA176) %>%
  t() 

# Make the circular plot
chordDiagram(chord_data, transparency = 0.5)

# Clockwise labels may require increasing the canvas margins
circos.par(canvas.xlim=c(-1.5,1.5),canvas.ylim=c(-1.5,1.5))
#df.test=df[,c(1,2)]

set.seed(2)

circos.clear()

par(cex = 0.5)

grid.col = c(
  AA171 = 'blue4',
  AA297 = "blue2",
  AA337 = "blue3",
  AA324 = "lightblue4",
  AA735 = "grey",
  AA315 = "grey50",
  AA174 = "black",
  AE638 = "pink",
  AA179 = "red1",
  AA175 = "red2",
  AA176 = "red3",
  AA747 = "red4"
)

chordDiagram(chord_data, annotationTrack = "grid", preAllocateTracks = 1, grid.col = grid.col)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  xlim = get.cell.meta.data("xlim")
  ylim = get.cell.meta.data("ylim")
  sector.name = get.cell.meta.data("sector.index")
  circos.text(mean(xlim), ylim[1] + cm_h(1.5), sector.name, facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5), cex =1.5)
  circos.axis(h = "top", labels.cex =1, major.tick.percentage = 0.2, sector.index = sector.name, track.index = 2)
}, bg.border = NA)
```


```{r}
clust_id = "AA176"

detected_clusters %>% filter(primary_cluster_id == clust_id) %>% arrange(desc(size)) %>% group_by(secondary_cluster_id) %>% add_count() %>% slice_head(n = 1) %>% arrange(desc(n))


set.seed(1)

genometa %>% select(name, ST_new) %>% full_join(mobtyper_df, by = 'name') %>% filter(secondary_cluster_id %in% c("AH856", "AH858", "AH855")) %>% group_by(secondary_cluster_id, ST_new) %>% add_count() %>% arrange(desc(n)) %>% slice_sample(n = 1) %>% select(name, ST_new, primary_cluster_id, secondary_cluster_id, n) %>% arrange(secondary_cluster_id, desc(n)) %>% group_by(secondary_cluster_id) %>% slice_head(n = 5)
```


```{r}
clust_id = "AA337"

detected_clusters %>% filter(primary_cluster_id == clust_id) %>% arrange(desc(size)) %>% group_by(secondary_cluster_id) %>% add_count() %>% slice_head(n = 1) %>% arrange(desc(n))

set.seed(1)

genometa %>% select(name, ST_new) %>% full_join(mobtyper_df, by = 'name') %>% filter(secondary_cluster_id %in% c("AI278", "AI277")) %>% group_by(secondary_cluster_id, ST_new) %>% add_count() %>% arrange(desc(n)) %>% slice_sample(n = 1) %>% select(name, ST_new, primary_cluster_id, secondary_cluster_id, n) %>% arrange(secondary_cluster_id, desc(n)) %>% group_by(secondary_cluster_id) %>% slice_head(n = 5)
```

```{r}
clust_id = "AA171"

detected_clusters %>% filter(primary_cluster_id == clust_id) %>% arrange(desc(size)) %>% group_by(secondary_cluster_id) %>% add_count() %>% slice_head(n = 1) %>% arrange(desc(n))

set.seed(1)

genometa %>% select(name, ST_new) %>% full_join(mobtyper_df, by = 'name') %>% filter(secondary_cluster_id %in% c("AH824", "AH825", "AH826", "AH827", "AH828", "AH829")) %>% group_by(secondary_cluster_id, ST_new) %>% add_count() %>% arrange(desc(n)) %>% slice_sample(n = 1) %>% select(name, ST_new, primary_cluster_id, secondary_cluster_id, n) %>% arrange(secondary_cluster_id, desc(n)) %>% group_by(secondary_cluster_id) %>% slice_head(n = 5)
```

```{r}
#Write supp table 2
detected_clusters %>% write_delim("supplementary/Supplementary_Table_2.txt", delim = "\t")
                                                                                                                                                             
#Write supp table 1
mobtyper_df_wide <- mobtyper_df %>% mutate(primary_cluster_id = str_replace(pattern = "novel.*", replacement = "Novel (Unclustered)", primary_cluster_id)) %>% select(name, primary_cluster_id) %>% mutate(val_col = 1) %>% pivot_wider(names_from = "primary_cluster_id", values_from = val_col, values_fill = 0, values_fn = sum) %>% select(everything(), `Novel (Unclustered)`)

Supp_table_1 <- read_delim("supplementary/Supplementary_Table_1_base.txt", delim = "\t") %>% select(1:16)

Supp_table_1_extra_cols <- left_join(genometa, mobtyper_df_wide, by = 'name') %>% select(c(1:7, 56:ncol(.)))

Supp_table_1 <- Supp_table_1 %>% left_join(Supp_table_1_extra_cols, by = 'name') %>% select(-`PopPunk Cluster (Clusters merged by PopPunk are separated by underscores)`, -`cgMLST cluster (40 allelic distance)`)

Supp_table_1 %>% write_delim("supplementary/Supplementary_Table_1.txt", delim = "\t")
```


### Exploratory Heatmaps
```{r}
mobsuite_contigs <- read_delim("analysis/mobsuite_concatenated_mge_report.txt", delim = "\t")

mobsuite_contigs <- mobsuite_contigs %>% mutate(name_contig = paste0(sample_id, "_", str_replace(contig_id, " .*", "")))

# Conditionally open analysis/genotype.txt or unzip analysis/genotype.txt
if (file.exists("analysis/genotype.txt")) {
  abricate_data <- read_delim("analysis/genotype.txt", delim = "\t")
} else if (file.exists("analysis/genotype.txt.gz")) {
  abricate_data_file <- unzip("analysis/genotype.txt.gz")
  abricate_data <- read_delim(abricate_data_file, delim = "\t")
} else {
  stop("Neither analysis/genotype.txt nor analysis/genotype.txt.gz exists. Pull the repo again!")
}

abricate_data <- abricate_data %>% mutate(name_contig = paste0(name, "_", SEQUENCE))

mobsuite_contigs <- full_join(mobsuite_contigs, abricate_data, by = "name_contig")

mobsuite_contigs <- mobsuite_contigs %>% filter(molecule_type != "chromosome")

mobsuite_contigs <- mobsuite_contigs %>% filter(`%COVERAGE` >= 90, `%IDENTITY` >= 90)

mobsuite_contigs %>% select(name, primary_cluster_id, GENE, DATABASE) %>% mutate(GENE_DB = paste0(DATABASE, "_", GENE)) %>% select(-GENE, -DATABASE) %>% mutate(val_col = 1) %>% pivot_wider(values_from = val_col, names_from = "GENE_DB", values_fill = 0, values_fn = sum)

pheatmap1_df <- mobsuite_contigs %>% select(name, primary_cluster_id, GENE, DATABASE) %>% mutate(GENE_DB = paste0(DATABASE, "_", GENE)) %>% select(-GENE, -DATABASE) %>% mutate(val_col = 1) %>% pivot_wider(values_from = val_col, names_from = "GENE_DB", values_fill = 0, values_fn = sum) %>% filter(primary_cluster_id %in% base_table$primary_cluster_id[1:12]) %>% select(-name) %>% group_by(primary_cluster_id) %>% summarise(across(everything(), ~ sum(. > 0, na.rm = TRUE) / n(), .names = "prop_{col}")) %>% column_to_rownames('primary_cluster_id') %>% select_if(~ sum(.) != 0) %>% select(matches("EC_custom|vfdb"))

pheatmap1_df %>% pheatmap(fontsize_col = 4, cluster_cols = F)

cols_heatmap <- colnames(pheatmap1_df) %>% str_replace(., "prop_", "ABRICATE..")

pheatmap2_df <- mobtyper_df %>%
    # Select relevant columns from mobtyper_df
    select(name, primary_cluster_id) %>%
    # Join with genometa on 'name'
    left_join(genometa, by = 'name') %>%
    # Select relevant columns after the join
    select(name, primary_cluster_id, matches("ABRICATE..vfdb"), matches("ABRICATE..EC_custom")) %>%
    # Inner join with base_table
    inner_join(base_table, by = "primary_cluster_id") %>%
    # Select columns that are also in detected_clusters
    select(name, primary_cluster_id, colnames(.)[colnames(.) %in% cols_heatmap]) %>%
    # Group by primary_cluster_id
    group_by(primary_cluster_id) %>%
    # Calculate the proportion of rows carrying each gene
    summarise(across(matches("ABRICATE"), ~ sum(. > 0, na.rm = TRUE) / n(), .names = "prop_{col}")) %>%
    # Ensure unique rows for each primary_cluster_id
    unique() %>%
  # Filter only for clusters from our table
  filter(primary_cluster_id %in% base_table$primary_cluster_id[1:12]) %>%
    # Set primary_cluster_id as row names
    column_to_rownames("primary_cluster_id") %>%
      # Remove columns that are all zeros
    select_if(~ !(all(. == 0, na.rm = TRUE))) %>%
  rename_with(~ gsub("ABRICATE..", "", .), everything())

pheatmap2_df <- pheatmap2_df %>% select(colnames(pheatmap1_df))



```
```{r}
pheatmap3_df <- detected_clusters %>%
  #rename sample_id
  rename('name' = sample_id) %>%
    select(name, primary_cluster_id, matches("ABRICATE..vfdb"), matches("ABRICATE..EC_custom")) %>%
    # Inner join with base_table
    inner_join(base_table, by = "primary_cluster_id") %>%
    # Select columns that are also in detected_clusters
    select(name, primary_cluster_id, colnames(.)[colnames(.) %in% cols_heatmap]) %>%
    # Group by primary_cluster_id
    group_by(primary_cluster_id) %>%
    # Calculate the proportion of rows carrying each gene
    summarise(across(matches("ABRICATE"), ~ sum(. > 0, na.rm = TRUE) / n(), .names = "prop_{col}")) %>%
    # Ensure unique rows for each primary_cluster_id
    unique() %>%
  # Filter only for clusters from our table
  filter(primary_cluster_id %in% base_table$primary_cluster_id[1:12]) %>%
    # Set primary_cluster_id as row names
    column_to_rownames("primary_cluster_id") %>%
      # Remove columns that are all zeros
    select_if(~ !(all(. == 0, na.rm = TRUE))) %>%
  rename_with(~ gsub("ABRICATE..", "", .), everything())

pheatmap3_df %>%
    # Generate the heatmap with specified font sizes
    pheatmap(fontsize_col = 4, cluster_cols = F)

pheatmap2_df <- pheatmap2_df %>% select(colnames(pheatmap3_df))

pheatmap2_df %>%
    # Generate the heatmap with specified font sizes
    pheatmap(fontsize_col = 4, cluster_cols = F)

pheatmap3_df - pheatmap2_df

pheatmap1_df %>% select(colnames(pheatmap3_df)) %>%
    # Generate the heatmap with specified font sizes
    pheatmap(fontsize_col = 4, cluster_cols = F)
```


```{r}
df3 <- pheatmap2_df %>% rownames_to_column('rownem') %>% mutate(rownem = str_replace(rownem, '$',"_")) %>% column_to_rownames('rownem') %>% bind_rows(pheatmap3_df) %>% rownames_to_column('rownem') %>% arrange(rownem) %>% column_to_rownames('rownem') 

df3 %>% pheatmap(cluster_rows = F, colorRampPalette(c("white", "red"))(100))

pheatmap1_df %>% select(colnames(df3)) %>% rownames_to_column('rownem') %>% mutate(rownem = str_replace(rownem, '$',"__")) %>% column_to_rownames('rownem') %>% bind_rows(df3) %>% rownames_to_column('rownem') %>% arrange(rownem) %>% column_to_rownames('rownem') %>% pheatmap(cluster_rows = F , colorRampPalette(c("white", "red"))(100))
```

```{r}
pheatmap4_df <- detected_clusters %>%
  #rename sample_id
  rename('name' = sample_id) %>%
    select(name, primary_cluster_id, matches("ABRICATE..vfdb"), matches("ABRICATE..EC_custom"), matches("ABRICATE..CARD")) %>%
    # Inner join with base_table
    inner_join(base_table, by = "primary_cluster_id") %>%
    # Select columns that are also in detected_clusters
    select(name, primary_cluster_id, matches("ABRICATE")) %>%#, colnames(.)[colnames(.) %in% cols_heatmap]) %>%
    # Group by primary_cluster_id
    group_by(primary_cluster_id) %>%
    # Calculate the proportion of rows carrying each gene
    summarise(across(matches("ABRICATE"), ~ sum(. > 0, na.rm = TRUE) / n(), .names = "prop_{col}")) %>%
    # Ensure unique rows for each primary_cluster_id
    unique() %>%
  # Filter only for clusters from our table
  filter(primary_cluster_id %in% base_table$primary_cluster_id[1:12]) %>%
    # Set primary_cluster_id as row names
    column_to_rownames("primary_cluster_id") %>%
      # Remove columns that are all zeros
    select_if(~ !(all(. == 0, na.rm = TRUE))) %>%
  rename_with(~ gsub("ABRICATE..", "", .), everything())

pheatmap4_df %>%
    # Generate the heatmap with specified font sizes
    pheatmap(fontsize_col = 4, cluster_cols = F)
```

```{r}
strains_w_major_plasmids <- mobtyper_df %>% filter(primary_cluster_id %in% c("AA171",
"AA176",
"AA337",
"AA352",
"AA174",
"AA324",
"AA175",
"AA297",
"AA315",
"AE638",
"AA179",
"AA735")) %>% select(name, primary_cluster_id)

strains_w_major_plasmids_genometa <- genometa %>% filter(name %in% strains_w_major_plasmids$name)

strains_w_major_plasmids_genometa <- full_join(strains_w_major_plasmids_genometa, strains_w_major_plasmids, by = 'name')

table_for_genes <- strains_w_major_plasmids_genometa %>%
    group_by(primary_cluster_id) %>%
    summarise(
      "ColV" = paste0(sum(ColV > 0)," (", scales::percent(mean(`ColV`), accuracy = 0.1), ")"),
      "senB" = paste0(sum(`ABRICATE..vfdb_senB` > 0)," (", scales::percent(mean(`ABRICATE..vfdb_senB`), accuracy = 0.1), ")"),
      "HPI Carriage" = paste0(sum(`ABRICATE..vfdb_fyuA/psn` > 0)," (", scales::percent(mean(`ABRICATE..vfdb_fyuA/psn`), accuracy = 0.1), ")"),
              "CIA Status" = paste0(sum(`Strain_CIA_status` > 0)," (", scales::percent(mean(`Strain_CIA_status`), accuracy = 0.1), ")"),
              "MDR Status" = paste0(sum(`Strain_MDR_status` > 0)," (", scales::percent(mean(`Strain_MDR_status`), accuracy = 0.1), ")"),
              "intI1 Carriage" = paste0(sum(`ABRICATE..EC_custom_intI1_HQ730118.1` > 0)," (", scales::percent(mean(`ABRICATE..EC_custom_intI1_HQ730118.1`), accuracy = 0.1), ")")) %>%
    mutate("Plasmid Type" = case_when(primary_cluster_id == "AA171" ~ 'senB',
                                      primary_cluster_id == "AA297" ~ "senB",
                                      primary_cluster_id == "AA337" ~ "senB",
                                      primary_cluster_id == "AA324" ~ "senB",
                                      primary_cluster_id == "AA352" ~ "senB",
                                      primary_cluster_id == "AA735" ~ "Other",
                                      primary_cluster_id == "AA315" ~ "Other",
                                      primary_cluster_id == "AA174" ~ "Other",
                                      primary_cluster_id == "AE638" ~ "ColV",
                                      primary_cluster_id == "AA179" ~ "ColV",
                                      primary_cluster_id == "AA175" ~ "ColV",
                                      primary_cluster_id == "AA176" ~ "ColV",
                                      primary_cluster_id == "AA747" ~ "ColV")) %>%
    select(primary_cluster_id, `Plasmid Type`, everything())

tmp_df <- as.data.frame(
  c("AA171",
"AA176",
"AA337",
"AA352",
"AA174",
"AA324",
"AA175",
"AA297",
"AA315",
"AE638",
"AA179",
"AA735"))



colnames(tmp_df) <- 'primary_cluster_id'

left_join(tmp_df, table_for_genes)
```

