---
title: "AusTrakka TEA Quality Control"
author: "Max Cummins"
date: "`r Sys.Date()`"
output:
    html_document:
        theme: spacelab
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Read in our packages
library(tidyverse)
library(ggplot2)
library(xml2)

#Define our not in function
`%nin%` <- Negate(`%in%`)
```

# Quality control AusTrakka Genomes

Here we perform quality control analysis on the cohort.

Our quality control metrics are as follows:

```{r set_QC_params, echo=FALSE}
### QC Metrics

##Assembly stats
        #Max number of scaffolds
QC_max_scaff <- 800
        #Minimum N50
QC_min_N50 <- 15000

#QC MLST metrics
        #Intended MLST scheme as per mlst (https://github.com/tseemann/mlst)
        #One of the following (column 1):
                #https://github.com/tseemann/mlst/blob/master/db/scheme_species_map.tab
QC_ST_scheme <- "ecoli"

#Bracken metrics
        #Expected species ID as per Bracken. Latin name should work
QC_expected_species <- "Escherichia coli"

#CheckM metrics
        #Minimum required 'Completion' metric
QC_min_complete <- 95
        #Maximum 'Contamination' metric
QC_max_contamination <- 5

#QC working directory
QC_report_path <- "analysis/QC_report.txt"

#MLST path
mlst_path <- "analysis/mlst.txt"
```

## Download our expected genome sizes

Next we need to determine our expected genome size for our samples. For that we query NCBI to determine the min and max genome size in the database.

Below is based on *E. coli*, but you can change the taxid to that of another species and it will still work.

```{r download_expected_genome_size, echo=TRUE}
#Download an XML file containing expected genome size for E. coli
#Change the taxid value for genome sizes for a different species
if(!file.exists("delims/expected_genome_size.xml")){
        download.file("https://api.ncbi.nlm.nih.gov/genome/v0/expected_genome_size/expected_genome_size?species_taxid=562&length=1", 'delims/expected_genome_size.xml')
        }

#Read in our XML file
expected_genome_sizes <- read_xml(x = "delims/expected_genome_size.xml")

#Pull out the minimum genome size and allow a buffer of 5%
#i.e. we allow anything within 5% of the smallest genome for E. coli in NCBI DBs
min_genome_size <- xml_text(xml_find_first(expected_genome_sizes, 'minimum_ungapped_length')) %>% as.numeric() * 0.95

#Pull out the minimum genome size and allow a buffer of 5%
#i.e. we allow anything within 5% of the largest genome for E. coli in NCBI DBs
max_genome_size <- xml_text(xml_find_first(expected_genome_sizes, 'maximum_ungapped_length')) %>% as.numeric() * 1.05
```

## Create temporary composite QC report

Until we have a proper QC file with everything included we first need to make a composite QC file from the ones we have for the TEA genomes and the original AusTrakka cohort

```{r create_temporary_QC_file, eval=FALSE, include=FALSE}

#Read in our TEA genome QC report
QC_report_TEA <- read_delim(QC_report_path, 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE,
    show_col_types = FALSE)

#Read in our original genome QC report
QC_report_original <- read_delim("analysis/QC_report.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE,
    show_col_types = FALSE)

#Bind our tables
QC_report <- bind_rows(QC_report_TEA, QC_report_original)

#Read in our TEA genome MLST file
mlst_TEA <- read_delim(mlst_path, 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE,
    show_col_types = FALSE)

#Read in original genome MLST file
mlst_original <- read_delim("analysis/mlst.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE,
    show_col_types = FALSE)

#Bind our tables
mlst <- bind_rows(mlst_TEA, mlst_original)

```

## Quality control filtering

Now we filter our genomes based on the metrics we provided above.

```{r simple_QC, echo=FALSE, include=FALSE}

#Read in QC report if it doesnt already exist in memory
if(!exists("QC_report")){
        QC_report <- read_delim(QC_report_path, 
                                delim = "\t", escape_double = FALSE, 
                                trim_ws = TRUE,
                                show_col_types = FALSE)
}

##Duplicate rows for the same samples
#Samples w multiple best matches occurring in the table twice
#Group by sample name and collapse tied best matches for kraken
#Remove duplicate rows
QC_report <- QC_report %>% group_by(name) %>% mutate(BRACKEN.species_second_best_match = toString(BRACKEN.species_second_best_match)) %>% unique()

#Read in MLST data
if(!exists("mlst")){
        mlst <- read_delim(mlst_path, 
                           delim = "\t", escape_double = FALSE, 
                           trim_ws = TRUE,
                           show_col_types = FALSE)
}

#Add a prefix for the MLST column
mlst <- mlst %>% rename_with(~gsub("^", "MLST.", .x)) %>% rename("name" = "MLST.name")

#Join our MLST and QC data
QC_report_w_mlst <- inner_join(QC_report, mlst, by = "name")

#Create a list of which genomes failed why
QC_report_pass_or_fail <- QC_report_w_mlst %>%
        group_by("Expected_species_ID" = BRACKEN.species_best_match == QC_expected_species,
                 "Expected_MLST_alleles" = MLST.scheme == QC_ST_scheme,
                 "Sufficient_completeness" = CHECKM.Completeness >= QC_min_complete,
                 "Sufficient_purity" = CHECKM.Contamination <= QC_max_contamination,
                 "Sufficient_scaffold_number" = ASSEMBLY_STATS.number <= QC_max_scaff,
                 "Sufficient_N50" = ASSEMBLY_STATS.N50 >= QC_min_N50,
                 "Within_expected_genome_size_range" = ASSEMBLY_STATS.total_length >= min_genome_size & ASSEMBLY_STATS.total_length <= max_genome_size)

#Create a list of QC passing genomes
QC_pass <- QC_report_pass_or_fail %>%
        filter(Expected_species_ID == TRUE,
               Expected_MLST_alleles == TRUE,
               Sufficient_completeness == TRUE,
               Sufficient_purity == TRUE,
               Sufficient_scaffold_number == TRUE,
               Sufficient_N50 == TRUE,
               Within_expected_genome_size_range == TRUE)
                     
#Create a list of QC failing genomes
QC_fail <- QC_report_pass_or_fail %>% filter(name %nin% QC_pass$name)

#Create a list of QC passing genomes
QC_pass_names <- QC_pass %>% as.data.frame() %>% select(name)
```

```{r, write_files}
#Check if our folder exists before continuing
if(!dir.exists("delims")){
        dir.create("delims")
}

#Check if our file exists before continuing
if(!file.exists("delims/QC_pass_names.txt")){
        
        #Save our list of QC_names
        write_delim(QC_pass_names, file = "delims/QC_pass_names.txt")
}

#Check if our file exists before continuing
if(!file.exists("delims/QC_data_QC_passes.csv")){
        #Save following files to computer
        write.csv(QC_pass,"delims/QC_data_QC_passes.csv", row.names = TRUE)
}

#Check if our file exists before continuing
if(!file.exists("delims/QC_data_QC_fails.csv")){
        #Save our QC data of genomes which failed QC to file
        write.csv(QC_fail,"delims/QC_data_QC_fails.csv", row.names = TRUE)
}

#Check if our file exists before continuing
if(!file.exists("delims/QC_report_pass_or_fail.csv")){
        #Save our report of QC passes or fails to file
        write.csv(QC_report_pass_or_fail,"delims/QC_report_pass_or_fail.csv", row.names = TRUE)
}
```

## Results



## Failed genomes

Below is a visualisation of the genomes which failed to meet our QC cut-offs.

```{r plot_failed_summary, echo=FALSE, fig.align="center", fig.width=12, fig.height=8}

library(ComplexUpset)

#Pull out the intersections we want to plot based on
reason_for_QC_fail <- QC_fail %>% select(21:27) %>% colnames()

#Generate out plot of genomes which failed QC
upset(QC_fail,
      reason_for_QC_fail,
      name = 'Reason for QC Failure',
      keep_empty_groups = TRUE,
      annotations = list(
        'Top Species (Bracken)'=(
            ggplot(mapping=aes(fill=BRACKEN.species_best_match))
            + geom_bar(stat='count', position='fill')
            + scale_y_continuous(labels=scales::percent_format())
            + ylab('Species')),
        'Top Species (MLST)'=(
            ggplot(mapping=aes(fill=MLST.scheme))
            + geom_bar(stat='count', position='fill')
            + scale_y_continuous(labels=scales::percent_format())
            + ylab('Species')),
        'Assembly Stats' =(
                 ggplot(mapping=aes(y=ASSEMBLY_STATS.total_length/1000000))
                + geom_point(na.rm=TRUE)
                + geom_violin(alpha=0.5, na.rm=TRUE)
                + geom_hline(yintercept=min_genome_size/1000000,
                             color = "#bebeda",
                             linetype = "dashed")
                + geom_hline(yintercept=max_genome_size/1000000,
                             color = "#bebeda",
                             linetype = "dashed")
                + ylab('Genome Size (Mbp)')     
            )
        )
    )
```
